---
title: Introduction to Fitting the Continuous Response Model (CRM) Using the EstCRM Package
subtitle:
author:
  name: Cengiz Zopluoglu
  affiliation: University of Oregon
date: 09/9/2024
output: 
  html_document:
    keep_md: false
    theme: journal
    highlight: haddock
    code_folding: show
    toc: yes
    toc_depth: 4
    toc_float: yes
always_allow_html: true
urlcolor: blue
mainfont: cochineal
sansfont: Fira Sans
monofont: Fira Code ## Although, see: https://tex.stackexchange.com/q/294362

## Automatically knit to both formats:
knit: (function(inputFile, encoding) {
 rmarkdown::render(inputFile, encoding = encoding, 
 output_format = 'all') 
 })
---

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position=c('top','right'),color='#33C1FF')
```


<style>
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    z-index: 2;
    color: #fff;
    background-color: #33C1FF;
    border-color: #97CAEF;
}

</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "",fig.align='center',message = FALSE,warning = FALSE)
require(here)
require(ggplot2)
require(plot3D)
require(kableExtra)
require(knitr)
require(giski)
require(magick)
options(scipen=99)

```

`r paste('[Updated:',format(Sys.time(),'%a, %b %d, %Y - %H:%M:%S'),']')`

<font color="black">

For a more detailed introduction to the Continuous Response Model (CRM), you can visit [this webpage](https://cengiz.me/posts/crm-stan/). The webpage includes a thorough explanation of the model, theoretical background, and step-by-step instructions for fitting the model using Stan. 

This tutorial will focus on demonstrating how to use the `EstCRM` package for parameter estimation.

# 1. Overview 

The `EstCRM` package provides functions to estimate parameters for the Samejima's Continuous Response Model (CRM) via marginal maximum likelihood estimation (MMLE) using the EM algorithm. The package allows users to estimate item and person parameters, evaluate item fit, draw 3D item response curves, and simulate data.

# 2. Installation and Setup

Before starting, make sure the `EstCRM` package is installed and loaded in your R environment. 

```{r}

# Install the package (if needed)
# install.packages("EstCRM")

# Load the package

require(EstCRM)

```

# 3. Estimating Item Parameters with the `EstCRMitem()` function

Let's start by estimating item parameters using the dataset EPIA, which contains five items scored on a continuous scale. The data is part of the package. This dataset came from a published study and was kindly provided by Dr. Ferrando. A group of 1,033 undergraduate students were asked to check on a 112 mm line segment with two end points (almost never, almost always) using their own judgement for the five items taken from the Spanish version of the EPI-A impulsivity subscale. The direct item score was the distance in mm of the check mark from the left end point ([Ferrando, 2002](https://doi.org/10.1207/s15327906mbr3704_05)).

```{r}

# Load the dataset
data(EPIA)
head(EPIA)

# Estimate the item parameters using the EM algorithm

item_params <- EstCRMitem(data        = EPIA, # name of the data object
                          max.item    = c(112, 112, 112, 112, 112), # maximum possible scores for each item
                          min.item    = c(0, 0, 0, 0, 0), # minimum possible scores for each item
                          max.EMCycle = 500, # maximum number of EM Cycles used in the iteration
                          converge    = 0.01) # a criteria value to stop the iterations
```

```{r}
# Extract the item parameter estimates

item_params$param
```

# 4. Estimating Person Parameters with the `EstCRMperson` function

After estimating item parameters, we can estimate person ability parameters (theta) using the `EstCRMperson` function. The person parameters can be computed once the item parameters are available.

```{r}

person_params <- EstCRMperson(data     = EPIA, 
                              ipar     = item_params$param, 
                              min.item = c(0, 0, 0, 0, 0), 
                              max.item = c(112, 112, 112, 112, 112))

head(person_params$thetas)

```

# 5. Item residuals using the `fitCRM` function

We can compute standardized item fit residual statistics using the `fitCRM` function. This helps evaluate how well the model fits the data.

```{r}

fit_results <- fitCRM(data       = EPIA, 
                      ipar       = item_params$param, 
                      est.thetas = person_params, 
                      max.item   = c(112, 112, 112, 112, 112), 
                      group      = 10)

fit_results$fit.stat
```


</font>


